### 2021-03-06 paper/paper.tex
paper: Improve results section
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

\documentclass[11pt,a4paper,english,twocolumn]{article}
\usepackage{unir_paper}
\usepackage{subcaption}
\usepackage{tablefootnote, multirow}

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Deep Embedded Clustering for Open Cluster Characterization with Gaia DR2 Data}
\author{CD. Álvaro, C. Guzman, J. Álvaro}
\date{24th of January, 2021}

%---------------------------
%marges
%---------------------------
%\usepackage[margin=1.9cm]{geometry}
%---------------------------
%---------------------------
%---------------------------
%---------------------------

\resumen{
Characterize and understand \emph{Open Clusters} (OCs) allow us to
understand better properties and mechanisms about the Universe such as stellar
formation and the regions where these events occur. They also provide information
about stellar processes and the evolution of the galactic disk.

In this paper, we present a novel method to characterize OCs. Our method employs a
model built on \emph{Artificial Neural Networks} (ANNs). More specifically, we adapted
a state of the art model, the \emph{Deep Embedded Clustering} (DEC) model for our purpose.
The developed method aims to improve classical state of the arts techniques. We improved
not only in terms of computational efficiency (with lower computational requirements),
but in usability (reducing the number of hyperparameters to get a good characterization
of the analyzed clusters). For our experiments, we used the \emph{Gaia DR2 database} as
the data source, and compared our model with the clustering technique \emph{K-Means}. Our
method achieves good results, becoming even better (in some of the cases) than current techniques.
}

\palabrasclave{
    characterization,
    data analysis,
    deep embedded clustering,
    gaia,
    machine learning,
    open cluster
}

\begin{document}
\twocolumn[
\begin{@twocolumnfalse}
\maketitle
\end{@twocolumnfalse}
]

%\renewcommand{\listfigurename}{Índice de Ilustraciones}
%\renewcommand{\listtablename}{Índice de Tablas}
%\renewcommand{\contentsname}{Índice de Contenidos}
%\renewcommand{\figurename}{Figura}
%\renewcommand{\tablename}{Tabla}
%\twocolumn

%\frontmatter
%\tableofcontents
%\listoffigures
%\listoftables

\section{Introduction}

Stellar OCs~\cite{janes1982open} are groups of stars gravitationally
bounded originated from a single molecular gas cloud.
They share the same chemical composition and age, and that is the reason
that their metallicity must be uniform since those stars were born from
the same gas cloud and at the same time stage. Likewise, they have similar
relative positions inherited from their original gas cloud, which means that
their distances to the Earth are the same. Therefore, they have a narrow
dispersion in their parallax value. Moreover, they share similar values
of proper motion, both in right ascension and declination. For all these
properties, we can see why the stellar OCs are relevant to understand the
spiral structure, dynamics, and chemical evolution of our galaxy.

The study of OCs advanced thanks to the huge and precise dataset
from the Gaia DR2 mission, available since 2018 from~\cite{collaboration2016description}
and~\cite{gaia2018gaia}. Although, the Gaia DR2 database misreported the chemical
composition of some OCs, the database has helped to review already known OCs and
to find new ones.

Characterize an OC refers to the statistical determination of the stars that
form it. Therefore, the characterization is based on the probability of a star
being a member of the cluster~\cite{sampedro2016caracterizacion}.

Normally, the characterization process considers astrometric features such as
proper motion in right ascension and declination, or parallax.
In some cases~\cite{oliveira2013fitting}, it also considers photometric features,
which helps to generate the H-R diagram~\cite{hypki2018gaia} of the OC candidate stars.
Those OC candidates should present a sharp profile corresponding to an isochrone curve
derived from a theoretical model. The model considers metallicity, mass and brightness
of the stars involved. For all this purpose, we used a set of tools that require
supervised and parameterized models. It means that we need a previous knowledge of the
cluster, otherwise we must repeat the process iteratively to achieve valid results.

In the present paper, we propose a machine learning model capable of characterizing
OCs inside a stellar region with no previous knowledge about the region.
Our model takes advantage of some features (such as proper motion in right ascension
and declination, and parallax) to train an ANN.
The ANN clusters all the stars within a given region in groups.
One of those groups will be the open cluster that interests us.

Our model contributes with several novelties: i) it is \emph{non-supervised} and
\emph{non-parameterized}, making easier the automation process of analyzing a
wide range of regions with different typologies, and ii) it is computationally
efficient to run in common workstations because it has been developed with
Python using the Keras framework which takes advantage of modern GPUs to perform
its computations. That increases significantly the computational capacity of
our model doing it possible to run on regular workstations.

This paper is organized as follows. Section~\ref{sec:related_works} presents
some related works and Section~\ref{sec:aims_and_method} describes in detail
our model. The following Section~\ref{sec:results}, shows the results and compares
them with a method that involves tools from the Virtual Observatory (VO) such as
Clusterix~\cite{balaguer2020clusterix}. Finally, the last section concludes and
outlines future research lines.

\section{Related Works}
\label{sec:related_works}

In~\cite{castro2020hunting}, they present one of the initial approaches to detect OCs.
They search for overdensities in the astrometric space of the galactic disk. Then,
they identify with photometric information possible OCs. Although it seems easy,
it is quite hard to face the problem that the near field around the OC has two kinds of
populations. The first population belongs to the members of the OC stars (from tens or
hundreds to a few thousand). The second population has a background of stars that do
not belong to the OC (from tens to hundreds of thousands). Thus, we summarize our problem
in finding out which stars belong to the OC.

Other works (TOPCAT~\cite{taylor2005topcat}, Clusterix 2.0~\cite{balaguer2020clusterix},
Aladin~\cite{bonnarel2000aladin}, or VOSA~\cite{bayo2008vosa}), first analyze the proper
motion configuration space of the region of interest. TOPCAT cannot find an open cluster.
Instead, it requires parameterizing some of the properties of the cluster. This last
process requires previous knowledge of the cluster. In other words, they perform a supervised
and manual selection of groups based on overdensities in the proper motion configuration space.
Clusterix 2.0 is an interactive web-based tool that can help the choice of groups. Take the
proper motion diagram without making any previous assumptions about the membership of the
candidate star. And empirically determine the frequency functions. It employs normal Gaussian
kernel functions, defined as:

\begin{equation*}
  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ - \frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{j} \right)^{2}}{ h^{2}} \right]}
\end{equation*}

where \(( a, b )\) are the proper motion configuration space, \(( i, j )\) is a point
located in the center of that provides the maximum contribution for calculating the
local density, and \(h\) is the \emph{smoothing} parameter, which it is measured in the
same units as the proper motion.

Clusterix depends fundamentally on the selection of three radii in the studied region.
The inner radius contains stars that belong or not to the cluster. While the outermost
radii define a ring that only contains stars that do not belong to the cluster.
Furthermore, it is also sensitive to other parameters, such as the soften parameter
\(h\), or other restrictions related to the searched proper motions.

If the analysis performed by Clusterix is successful, it returns the probability that
each star belongs to the open cluster. The results can then be imported into TOPCAT
to continue the refinement process and obtain a valid characterization.

Another recent approach~\cite{castro2020hunting} solves the problem using machine
learning techniques. It searches systematically for overdensities in the astrometric
space of the galactic disk and subsequent identification of open clusters using
photometric information. It includes two phases:

\begin{enumerate}
  \item it employs DBSCAN, an unsupervised clustering algorithm, to search for
        overdensities,
  \item and applies a deep learning ANN to identify isochrone patterns within
        the detected overdensities and thus proceed to confirm them as OCs. They
        previously trained the ANN with magnitude diagrams.
\end{enumerate}

They carry out their experiments at the Barcelona Supercomputing Center, MareNostrum 42.
Thereby, the neural network handles the image recognition process with isochrone
patterns without applying theoretical models derived from values such as metallicity
or masses, among others. Their work concludes with the identification of 582 new open
clusters distributed along with the galactic disk for a galactic declination below 20
degrees, increasing the number of known open clusters by 45\%.

In contrast to~\cite{castro2020hunting}, we lack a supercomputer like MareNostrum
for our experiments. For this reason, our goal is to obtain a new novel method that
allows the characterization of open clusters without doing a blind search for
characterization clusters. We advocate the idea of building an unsupervised and
non-parameterized method. Therefore, it can be suitable for automated processes.

\section{Open Cluster Characterization Method}
\label{sec:aims_and_method}

In the literature, we have several clustering algorithms, such as K-Means, Mean-Shift
Clustering, DBSCAN~\cite{ester1996density} among others, that can help us achieve
our aim. Each one behaves better according to the distribution of the objects to
be grouped. However, we need to establish a large number of clusters to find the open cluster
we are looking for. It complicates the identification of the OC because most of the
time falls in many outliers.

The unsupervised \emph{Deep Embedding for Clustering Analysis} (\emph{DEC})
model~\cite{xie2016unsupervised} is a refinement K-Means based on ANN. It starts with
K-Means. And then, it trains an autoencoder to reduce the feature space and pass them
through a Clustering Layer which refines the previous selection.

With all this in mind, our method is based on the DEC model. We adapted it to manage data
groups based on the dynamic properties of the stars. Thus, we have an unsupervised
clustering model for open cluster characterization. The model fits a wide range of
clusters without the need for fine-tuning a high number of hyperparameters.

\subsection{Selection of the Catalog}

In this work, we make use of Gaia DR2 as DR3 has not been released in time for us to
include it. Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission
(located at L2, 1.5 million kilometers from Earth) and operational since 2014. The
catalog has high precision and accuracy astrometric data for more than 1.7 billion
stellar sources and magnitudes in three photometric filters (G, BP, and RP) for more
than 1,300 million sources.

We start by selecting a region from the OpenClust~\cite{dias2002new} catalog and
downloading it from the Gaia DR2 database. The radius of the downloaded region from Gaia
is 1.5 times greater than that recorded in OpenClust, ensuring that it includes several
stars that do not belong to the open cluster.

\subsection{Selection of Features}
\label{sec:feature_selection}

For the selection of features, we studied the Melotte 22 dataset.
Table~\ref{tab:sample_features_melotte22_region} describes an example of the region Melotte 22.
Each row represents the features of a star.

\begin{table}[!htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{c|c|c}
        % \textbf{source\_id} &
        \emph{pmra \((mas \cdot yr^{-1})\)} & \emph{pmdec \((mas \cdot yr^{-1})\)} & \emph{parallax \((mas)\)} \\
        \hline
        % 64035217900973056 &
        2.848682 & -3.291204 & 0.399680 \\
        % 64035217900973440 &
        0.894901 & -3.445501 & 0.416639 \\
        % 64035217901215616 &
        7.924372 & -0.241281 & 0.397743 \\
        % 64035351043566208 &
        -4.433802 & -2.584965 & 0.410695 \\
        % 64035385403304960 &
        0.055990 & -1.760018 & 0.413813 \\
      \end{tabular}
    }
    \caption{Features of Melotte 22 region.}
    \label{tab:sample_features_melotte22_region}
  \end{center}
\end{table}

Proper motion in right ascension (column pmra) and declination (column pmdec) seems
like a natural choice since stars belonging to the same OC share a common motion vector.
Parallax (column parallax) is another relevant feature. It lets us know how far stars
are from us. Besides, since all stars within an open cluster were born from the same
dust cloud, they must all have similar parallax.

However, we are not going to use these raw features. Instead, we take a combination
of them. Figure~\ref{fig:features_melotte_22} shows a pairwise relationship of our
combination of features from Melotte 22.

\begin{figure*}[!hbt]
  \centering
  \includegraphics[scale=0.5]{../figures/melotte_22/features_melotte_22.png}
  \caption{Pairwise relationships among variables using Melotte 22 data}
  \label{fig:features_melotte_22}
\end{figure*}

First, we correct proper motion in right ascension and declination by dividing them by
the parallax (variables pmra\_corr and pmdec\_corr, respectively). That way, we normalize
these quantities and help our clustering models to improve their performance. The modulus
of the proper motion (variable pmmod in Figure~\ref{fig:features_melotte_22}) is another
computed property that we considered. We use it to relate both features and therefore
force our model to keep them tight.

\subsection{Deep Open Clustering of Stars}
\label{sec:deep_open_clustering}

In this section, we explain our model in more detail.

One of the main problems to define our model is that we do not have a labeled
dataset to train a supervised model. To solve it, we employ an unsupervised self-trained
model. Thus, we adapted the Unsupervised \emph{DEC} model~\cite{xie2016unsupervised}
to our requirements.

Like the DEC model, we have two main components:

\begin{itemize}
  \item \textbf{deep autoencoder:} it is a neural network that learns to copy its input
        to its output. It is composed of encoder layers followed by decoder layers.
        We trained it before passing the data through the next component, the
        \emph{clustering layer}. Recent research~\cite{vincent2010stacked,hinton2006reducing}
        showed that the autoencoder provides meaningful and well-separated representations
        on real-world datasets.
  \item \textbf{clustering layer:} it receives data transformed by the autoencoder
        and iteratively is trained until a convergence criterion is met.
\end{itemize}

The deep autoencoder is used to transform the input data into a latent space using a
non-linear mapping function \(f_{\theta} : X \rightarrow Z\).

We trained the autoencoder before fitting the model to generate predictions.
Then, we use the encoder layers of the autoencoder with the aim of transforming
input data to the latent space \(Z\). Once the data has been transformed, a K-Means
clusterer is used to make an initial clustering. K-Means cluster centers are used
as the initial weights for the clustering layer.

With that initial configuration, the model iterates alternating between computing an
auxiliary target distribution (soft assignment) and minimizing the Kullback-Leiber (KL)
divergence~\cite{kullback1951information} to it. This unsupervised algorithm allows us
to improve the clustering.

As it was described in Section~\ref{sec:feature_selection}, the number of features we
deal is not too large. However, this latent space helps us to start with a reduced number
of features and avoids the \emph{``curse of dimensionality''}~\cite{bellman1961curse}.

In the soft assignment stage, we used \emph{Student's t-distribution} as a kernel to
measure the similarity between the embedded points and the cluster centroid.
While in the KL divergence minimization the algorithm iteratively refines clusters by learning
from their high confidence assignments with the help of an auxiliary target distribution.
The model is trained by matching the soft assignment to the target distribution.
The choice of this target distribution is crucial for model's performance.

In this work, we took the target distribution~\cite{xie2016unsupervised}
defined in Equation~\ref{eq:student_tdistribution}.

\begin{equation}
  p_{ij} = \frac{q^{2}_{ij} / f_{j}}{\sum_{j'}q^{2}_{ij'}/f_{j'}}
  \label{eq:student_tdistribution}
\end{equation}

This target distribution has the following properties:

\begin{itemize}
  \item strengthen predictions (i.e., improve cluster purity),
  \item put more emphasis on data points assigned with high confidence,
  \item normalize loss contribution of each centroid to prevent large clusters
        from distorting the hidden feature space.
\end{itemize}

\begin{figure*}[!hbt]
  \centering
  \includegraphics[scale=0.8]{../figures/dec_diagram.pdf}
  \caption{Layer setup of our model}
  \label{fig:dec_model_setup}
\end{figure*}

Figure~\ref{fig:dec_model_setup} shows the layer setup of our model.
It is simpler than the used in~\cite{xie2016unsupervised} because we selected a smaller
number of features. Therefore, using the same configuration would result in a model so
powerful that it would incur overfitting issues that cannot make the correct predictions.

\begin{figure*}[!htbp]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
      \caption{N = 2}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
      \caption{N = 5}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
      \caption{N = 8}
  \end{subfigure}
  \caption{K-Means with Melotte 22}
  \label{fig:kmeans_comparisons_melotte_22}
\end{figure*}

Finally, we can refine this selection by filtering those stars that are below and above
the 0.10 and 0.90 quantiles for each group, respectively. That way we remove the most
doubtful values from the selection.

\paragraph{Discussion.}

%% Comparación con K-Means, DEC y DEC filter
\begin{figure*}[!htp]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_pm_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_hr_diagram_melotte_25.png}
      \caption{K-Means}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
      \caption{DECOCC}
  \end{subfigure}
  \medskip
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_filtered_melotte_25.png}
      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_filtered_melotte_25.png}
      \caption{DECOCC (filt.)}
  \end{subfigure}
  \caption{Evolution of Melotte 25 characterization}
  \label{fig:melotte_25_characterization_evolution}
\end{figure*}

Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm
like K-Means to find two clusters. One for the desired OC and another for stars that do not
belong to the OC. However, the idea is not completely right. This is due to the fact that OC's
stars are surrounded by other stars with possibly similar properties. Therefore, setting the
number of clusters to two is too low to separate them properly.

Figure~\ref{fig:kmeans_comparisons_melotte_22} shows an example. It describes the results of
applying different configurations of K-Means in Melotte 22, N=2, N=5 and N=8. For each configuration
of K-Means, we show the proper motion in right ascension and declination (figures in the top position),
and the parallax distribution (figures in the down position).

As we can see in the parallax distribution, lower values of the number of clusters (e.g., N=2) assign
many stars to the group \emph{g0} (purple) that should belong to group \emph{g1}. These badly assigned stars
are the ones that have the parallax far from the 7.3mas resonance. In contrast, large values of the
number of clusters allow us to isolate more accurately the resonance in parallax at \(\approx 7.3mas\).
For instance, at N=8, more groups are better compartmentalized. Group \emph{g2} would be the one corresponding
to the OC that we are looking for, and all the other groups are those that we would like to be a single
group. That is the stars that do not belong to the OC. The disadvantage is that K-Means with N=8 forms more
groups of stars, which complicates the task of finding the desired OC, since we would like to get just two
groups. Therefore, with a K-Means we have to find a way to set the right value for the number of clusters
to isolate the searched cluster without creating too many groups. To solve this issue, we can try to
estimate the best number of clusters by using the \emph{silhouette score}~\cite{rousseeuw1987silhouettes}.

K-Means does a good job making an initial clustering. However, too many clusters arise
from this characterization and the OC is still polluted with stars that do not belong to it.
Moreover, we would like to reduce the amount of clusters too.

\newpage

\section{Results}
\label{sec:results}

We performed two tested of our model:

\begin{enumerate}
  \item with the Melotte 25 dataset, which is one of the most difficult dataset to characterize OC
        by its distribution and extension, because it is located in a region with 400,000 stars.
  \item with a variety of clusters of different typologies.
\end{enumerate}

In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, in the both
tests we compared our model against a K-Means classification algorithm. In addition, we perform a
second execution only in our model where we filtered the quantiles in the Melotte 25 dataset.

Our DECOCC model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter
Notebooks~\cite{Kluyver2016jupyter}. All tests were run on an Apple Mac Pro Late 2013 with a
2.7GHz 12-Core Intel Xeon E5-2697v2, and 64GB RAM 1866MHz DDR3. For the GPU, we use a graphic
card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available
at \href{https://github.com/cdalvaro/machine-learning-master-thesis}
{https://github.com/cdalvaro/machine-learning-master-thesis}}.

\subsection{Melotte 25 dataset}

Figure~\ref{fig:melotte_25_characterization_evolution} shows the evolution of the characterization
of Melotte 25. The first column corresponds to the characterization obtained by applying only
K-Means (with N=10). The middle column shows the result of our model. And the last column describes
the characterization made by our model but filtering the quantiles lower than 0.10 and higher than
0.90. The top row shows the proper motion configuration space, while the bottom one shows the
Hertzsprung–Russell diagrams.

%% Comparación con K-Means, DEC y DEC filter

The clustering made by K-Means arises 9 groups, but most of them contain stars that do not belong
to the Melotte 25 OC. However, in our model, the stars regroup and although we have the same 9
groups of K-Means, almost all of the stars are assigned to a group that would contain the stars
that would not belong to the OC. The remaining groups contain either the OC we are looking for or
other possible new OCs inside the studied region. Moreover, if we exclude those stars that remain
below 0.10 and above 0.90 percentiles, and execute again this refinement in our model (third column
in Figure~\ref{fig:melotte_25_characterization_evolution}), we observe an improvement in the
results of our DECOCC model.

\begin{figure}[!hbt]
  % \centering
  % \begin{subfigure}[t]{0.30\textwidth}
  %   \centering
  %   \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[t]{0.30\textwidth}
    \centering
    \includegraphics[width=0.36\textwidth]{../figures/melotte_25/dec_parallax_melotte_25.png}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[t]{0.30\textwidth}
  %   \centering
  %   \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
  % \end{subfigure}
  % \caption{Melotte 25 characterization.
  %          From left to right: proper motion configuration space,
  %          parallax histogram and H-R diagram.}
  \caption{Parallax histogram of Melotte 25 characterization.}
  \label{fig:result_melotte_25_dec}
\end{figure}

The dataset Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in
declination with a radius of 165 arcmin. As it was mentioned before, this region contains more than
400,000 stars making it difficult to group stars of the same region. Our model labels Melotte 25 as
group \emph{g5}. Figure~\ref{fig:result_melotte_25_dec} shows the parallax histogram.

\begin{table}[!htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{l|c|c|c|c}
        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
        \hline
        Simbad & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
        % Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
        K-Means & 79.936 \( \pm \) 3.7 & -45.362 \( \pm \) 3.8 & 20.901 \( \pm \) 0.31 & 374 \\
        DECOCC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
        DECOCC (filt.) & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
      \end{tabular}
    }
    \caption{Summary for the Melotte 25 results.}
    \label{tab:results_melotte_25}
  \end{center}
\end{table}

Table~\ref{tab:results_melotte_25} summarizes the results for the Melotte 25 dataset. The first column
shows the name of the method. Second and third columns describe the proper motion in right ascension and
declination, respectively. The fourth column shows the parallax. And the last column presents the number
of stars corresponding to the OC. Columns from second to fourth have their respective deviations computed
as the standard error of the predicted group.

By comparing the results obtained with our DECOCC model, we can see that they are in range with those
extracted from Simbad astronomical database~\cite{wenger2000simbad}. While the results provided by K-Means
model are far from the official data. This comparison demonstrates that our model improves the results
obtained with a raw K-Means.

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{l|c}
      \textbf{Hyperparameter} & \textbf{Value} \\
      \hline
      Number of Clusters & 9 \\
      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
      Kernel Initializer Seed & 11 \\
      Quantil Threshold & 0.1 \\
    \end{tabular}
    \caption{Hyperparameters for Melotte 25.}
    \label{tab:hyperparameters_melotte_25}
  \end{center}
\end{table}

Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for the characterization. We used
them to fine-tuning our model and improve its accuracy. Although they depend on the studied region, none of
them are related with physical properties of the cluster, which makes it easy to automate the selection of
hyperparameters.

\subsection{Clusters with different typologies}

We perform now more exhaustive tests of our DECOCC model using open clusters with different typologies (e.g.,
large, small, more extensive, or more concentrated within the region to be analyzed):

\begin{itemize}
  \item NGC 2516, which has the proper motion center deviated from the origin but embedded within a large cloud
        of stars with similar proper motions.
  \item NGC 2632, which, in addition to Melotte 22, is a cluster whose proper motion center is not at the origin
        and has a well-separated parallax center.
  \item NGC 2682, which has the parallax centered within the region's Gaussian, making it difficult its detection.
        Moreover, same as NGC 2516, the proper motion center is deviated from the origin.
  \item Melotte 22, as well as Melotte 25, is a cluster closer to our galaxy. That causes its membership stars to
        be more scattered than previous clusters which are more compact.
\end{itemize}

Table~\ref{tab:clusters_summary} describes more specific information about the previous clusters (we also include
information about the Melotte 25 open cluster, which was previously analyzed). The first column shows the name of
the cluster. The remaining columns are the right ascension, declination, radius, and the number of stars of studied
clusters. The column \# stars correspond to those stars contained within a cone of center \((\alpha, \delta)\) and
radius the cluster's radius multiplied by a factor of 1.5.

\begin{table}[htbp]
  \begin{center}
    \resizebox{\columnwidth}{!}{
      \begin{tabular}{l|c|c|c|c}
        \textbf{OC} & \emph{\( \alpha \) J2000 \((degrees)\)} & \emph{\( \delta \) J2000 \((degrees)\)}
        & \emph{Radius \((arcmin)\)} & \emph{\# stars} \\
        \hline
        NGC 2516 & 119.517 & -60.753 & 15 & 12,869 \\
        NGC 2632 & 130.1 & 19.667 & 35 & 13,167 \\
        NGC 2682 & 132.825 & 11.8 & 12.5 & 2,839 \\
        Melotte 22 & 56.75 & 24.117 & 60 & 61,552 \\
        Melotte 25 & 66.725 & 15.867 & 165 & 433,996 \\
      \end{tabular}
    }
    \caption{Properties of clusters for further testing of our model.}
    \label{tab:clusters_summary}
  \end{center}
\end{table}

%% Tabla comparativa de todos los open clusters
\begin{table*}[!htb]
  \begin{center}
    % \resizebox{\columnwidth}{!}{
      \begin{tabular}{c|l|c|c|c|c}
         & Method & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
        \hline
        \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2516}}
        & Simbad & -4.6579 \( \pm \) 0.0075 & 11.1517 \( \pm \) 0.0075 & 2.4118 \( \pm \) 0.0006 & 1727 \\
        & K-Means & -4.344 \( \pm \) 0.14 & 9.507 \( \pm \) 0.19 & 2.268 \( \pm \) 0.01 & 1542 \\
        & DECOCC & -4.426 \( \pm \) 0.17 & 9.952 \( \pm \) 0.20 & 2.436 \( \pm \) 0.01 & 1532 \\
        & DECOCC (filt.) & -4.502 \( \pm \) 0.14 & 10.114 \( \pm \) 0.17 & 2.392 \( \pm \) 0.004 & 1072 \\\\

        % \caption{NGC 2632 results.}
        \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2632}} & Simbad & -36.047 \( \pm \) 0.110 & -12.917 \( \pm \) 0.066 & 5.371 \( \pm \) 0.003 & - \\
        & K-Means & -26.352 \( \pm \) 0.82 & -15.828 \( \pm \) 0.76 & 5.394 \( \pm \) 0.03 & 629 \\
        & DECOCC & -20.012 \( \pm \) 0.69 & -14.742 \( \pm \) 0.58 & 4.686 \( \pm \) 0.03 & 894 \\
        & DECOCC (filt.) & -21.571 \( \pm \) 0.74 & -14.234 \( \pm \) 0.61 & 4.719 \( \pm \) 0.03 & 714 \\\\

        % \caption{NGC 2682 results.}
         \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2682}} & Simbad & -10.9737 \( \pm \) 0.0064 & -2.9396 \( \pm \) 0.0063 & 1.1325 \( \pm \) 0.0011 & 1194 \\
        & K-Means & -8.616 \( \pm \) 0.15 & -3.710 \( \pm \) 0.16 & 1.196 \( \pm \) 0.01 & 1374 \\
        & DECOCC & -8.926 \( \pm \) 0.15 & -3.550 \( \pm \) 0.15 & 1.144 \( \pm \) 0.005 & 1238 \\
        & DECOCC (filt.) & -9.619 \( \pm \) 0.13 & -3.317 \( \pm \) 0.13 & 1.140 \( \pm \) 0.003 & 990 \\\\

        % \caption{Melotte 22 results.}
        \multirow{4}{*}{\rotatebox[origin=c]{90}{Melotte 22}} & Simbad & 19.997 \( \pm \) 0.127 & -45.548 \( \pm \) 0.101 & 7.364 \( \pm \) 0.005 & 1326 \\
        & K-Means & 20.25 \( \pm \) 0.95 & -38.01 \( \pm \) 1.08 & 7.23 \( \pm \) 0.06 & 1378 \\
        & DECOCC & 23.67 \( \pm \) 1.29 & -46.23 \( \pm \) 1.50 & 8.04 \( \pm \) 0.09 & 878 \\
        & DECOCC (filt.) & 19.50 \( \pm \) 0.41 & -44.23 \( \pm \) 0.39 & 7.42 \( \pm \) 0.005 & 438 \\

      \end{tabular}
    % }
    \caption{Summary of the results for the clusters with different typologies.}
    \label{tab:app_results_ngc_2516}
  \end{center}
\end{table*}

As we can see, all those clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number
of stars that belong to them. For instance, NGC 2682 has 3,000 stars while Melotte 25 is inside a region with more
than 400,000 stars.

Table~\ref{tab:app_results_ngc_2516} describes a summary of the results for the clusters with different typologies.
Again, our DECOCC model presents similar results from those extracted from the astronomical database Simbad. Except
for the open cluster NGC 2632. In this cluster, the K-Means presents better results than our model. And this is so
because the center of the proper motion is not at the origin and generates a more separate center of parallax, which
makes it difficult to characterize the open cluster in our model. However, our model presents more robust results in
more difficult open clusters, such as the NGC 2682 that has the parallax centered within the region's Gaussian.

%% comparativa de las figuras de los open clusters
\begin{figure*}[!hbt]
  \rotatebox[origin=c]{90}{{\bfseries NGC 2516}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2516} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2632}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2632} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
    \caption{K-Means}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
    \caption{DECOCC}
  \end{subfigure}%
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
    \caption{DECOCC (filt.)}
  \end{subfigure}
  \caption{Summary of the characterization for the clusters NGC 2516 and NGC 2632.}
  \label{fig:ngc_2516_ngc_2632_characterization_evolution}
\end{figure*}

\begin{figure*}[!hbt]
  \rotatebox[origin=c]{90}{{\bfseries NGC 2682}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_parallax_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_filtered_ngc_2682.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries NGC 2682} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_hr_diagram_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_ngc_2682.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_filtered_ngc_2682.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries Melotte 22}\strut}
  \rotatebox[origin=c]{90}{parallax \strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
  \end{subfigure}

  \rotatebox[origin=c]{90}{{\bfseries Melotte 22} \strut}
  \rotatebox[origin=c]{90}{hr diagram\strut}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
    \caption{K-Means}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
    \caption{DECOCC}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
    \caption{DECOCC (filt.)}
  \end{subfigure}
  \caption{Summary of the characterization for the clusters NGC 2682 and Melotte 22.}
  \label{fig:melotte_22_ngc_2682_characterization_evolution}
\end{figure*}

Figures~\ref{fig:ngc_2516_ngc_2632_characterization_evolution} and~\ref{fig:melotte_22_ngc_2682_characterization_evolution}
show the groups found for the three models. For each open cluster, we describe the parallax and hr diagram.

K-Means and DECOCC have labeled NGC 2516 as \emph{g3}. Although in general, groups between K-Means and our
model may not match.

%% tabla comparativa de los hyperparametros
\begin{table}[htbp]
  \begin{center}

    % \resizebox{\columnwidth}{!}{
        \begin{tabular}{l|l|c}
          & \textbf{Hyperparameter} & \textbf{Value} \\
          \hline
          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2516}} & Number of Clusters & 8 \\
          & Clustering Layer & \(\left[ 50, 50, 60 \right]\) \\
          & Kernel Initializer Seed & 2 \\
          & Quantil Threshold & 0.15 \\\\

          % \caption{NGC 2632 DEC hyperparameters.}
          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2632}} & Number of Clusters & 10 \\
          & Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
          & Kernel Initializer Seed & 10 \\
          & Quantil Threshold & 0.1 \\\\

         % \caption{NGC 2682 DEC hyperparameters.}
          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2682}} & Number of Clusters & 10 \\
          & Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
          & Kernel Initializer Seed & 0 \\
          & Quantil Threshold & 0.1 \\\\

          % \caption{Melotte 22 DEC hyperparameters.}
          \multirow{4}{*}{\rotatebox[origin=c]{90}{Melotte 22}} & Number of Clusters & 5 \\
          & Clustering Layer & \(\left[ 50, 50, 200 \right]\) \\
          & Kernel Initializer Seed & 11 \\
          & Quantil Threshold & 0.1 \\

        \end{tabular}
    % }
    \caption{Hyperparameters used by DECOCC model for the open clusters with different typologies.}
    \label{tab:app_hyperparameters_ngc_2516}
  \end{center}
\end{table}

Table~\ref{tab:app_hyperparameters_ngc_2516} shows the hyperparameters used for characterizing the
open clusters with our method.

In all cases, the DECOCC model has resolved properly the identity of the cluster and has characterized
the membership stars, giving compatible results with the ones obtained with classic procedures and VO
tools.

Compared with the mentioned tools, our model can be categorized as a non-parameterized model since it
does not depend on parameters referred to the cluster itself but only hyperparameters such as the
initial number of clusters to be found, or the structure of the Clustering Layer used by DEC model.

The proposed method is also non-supervised since we do not need to tell the model which stars belong
to the cluster or assist it while its training stage.

Only at the end of the characterization, a fine-tuning driven by the user, can be applied in order to
improve the selection by discarding those stars which fall outside the given quantiles.

In any case, we do not make assumptions about the cluster profile, something that we have to take into
account when using VO tools. This is another evidence that our model is non-parameterized.

\section{Conclusions}

From the results shown in Section~\ref{sec:results}, we can say that, in general,
our model works fine identifying and characterizing open clusters.
So we have succeeded building a model non-parameterized and non-supervised
for open cluster characterization. However, some improvements should be done
in order to improve its accuracy and precision.

Another achievement is that we have developed a model that does not require
complex hardware. Instead, it can be run on workstations with a common GPU,
making it accessible to be implemented in many research centers.

We have built a non-parameterized model since the user is not required to provide
any information about the cluster profile or its properties.
Nevertheless some tweaks can be done regarding the cluster, like changing the initial
number of clusters or modifying the clustering layer structure in order to improve
the final characterization.

Another key point in our model is the initializer kernel.
This kernel prepares data before passing it to the ANN
and the results may vary significantly depending on this kernel.
This is something to avoid, so an improvement to solve this issue is necessary
in order to have a reliable model which does not depend on the dataset order.

Despite that, comparing our method with Clusterix, we can still say that the
number of parameters (or hyperparameters) in our model is smaller and that
we do not need previous knowledge about the studied cluster.
That makes easier to test different configurations and automate the process.

Furthermore, our model is able to deal with large and open regions. As an example,
it has succeeded characterizing Melotte 25 (Hiades) OC. Also, the relative proximity
of this cluster, 45 parsec, makes it difficult to characterize. Clusterix fails
characterizing this cluster. In this case, our method proposes at least two more
non-catalogued clusters. The analysis of these new clusters brings out an interesting
stellar dynamics in the studied region. It could mean that there was a recent
interaction among different molecular clouds.

For all this, we consider that the presented method is good enough to be included
in the Virtual Observatory toolset. Only a few tweaks must be done in order to make
our model compatible with other tools of the VO.

\subsection{Future Works}

One open point to do from now is testing the model with a wider range of clusters.
In general we could think of applying it to the whole VizieR catalogue.
That way we could have a better idea about the current limitations of our model
and we could make some adjustments to improve it. We could determine with higher
accuracy the typologies that our model works well with, and which ones the model
does not resolve properly. Furthermore, we could stablish different sets of
hyperparameters regarding the typology of the cluster.

With our model, new uncatalogued clusters arise from the data apart of the main ones.
Hence, it is also interesting the individual study of these new clusters.

Another possible change in our model would be using DBSCAN instead of K-Means
as our initial clustering algorithm. Maybe this algorithm gives us a better starting
point for the DEC model which could improve the results.

Finally, we have used the Gaia DR2 database as our data source for this work,
but recently the DR3 dataset has been released.
Therefore, it is evident the interest on testing our model with this new data.

\renewcommand{\refname}{References}
\bibliographystyle{unsrt}
\bibliography{references}

\end{document}